---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# saeHB.Spatial.Beta

We designed this package to provide several functions for area-level small area estimation under Spatial Simultaneous Autoregressive (SAR) and Leroux Conditional Autoregressive (CAR) models, accommodating survey design effect (DEFF) adjustments, using hierarchical Bayesian (HB) method with Beta distribution for variables of interest. Some datasets simulated by a data generation are also provided. The `rjags` package is employed to obtain parameter estimates using Gibbs Sampling algorithm. Model-based estimators involve the HB estimators which include the mean estimation, the estimated model coefficients, the random effect, and the random effect variance. For the reference, see Rao and Molina (2015), Kubacki and Jedrzejczak (2016), Leroux et al. (2000), and Chung and Datta (2020).

## Author
Boby Iwan, Cucu Sumarni

## Maintainer
Boby Iwan <bobyiwanboby2122@gmail.com>

## Functions
* `betadeff_sar()` Estimates small area means using a Spatial SAR Model under a Beta distribution, incorporating survey design effect (DEFF) adjustments.
* `beta_sar()` Estimates small area means using a Spatial SAR Model under a Beta distribution without DEFF adjustments, by estimating the unknown precision parameter.
* `betadeff_lerouxcar()` Estimates small area means using a Spatial Leroux CAR Model under a Beta distribution, incorporating survey design effect (DEFF) adjustments.
* `beta_lerouxcar()` Estimates small area means using a Spatial Leroux CAR Model under a Beta distribution without DEFF adjustments, by estimating the unknown precision parameter.
* `betadeff_nonspatial()` Estimates small area means using a Non-Spatial Model under a Beta distribution with Independent and Identically Distributed (IID) random effects, incorporating DEFF adjustments.
* `beta_nonspatial()` Estimates small area means using a Non-Spatial Model under a Beta distribution with IID random effects without DEFF adjustments, by estimating the unknown precision parameter.
* `build_w()` A utility function to construct spatial weights matrices (contiguity, distance, or kernel) required for spatial modeling.
* `moran_test()` A diagnostic function to perform Moran's I test for spatial autocorrelation.

## Installation
You can install the development version of saeHB.Spatial.Beta from [GitHub](https://github.com/BobyIwan/saeHB.Spatial.Beta) with:

``` r
# install.packages("devtools")
devtools::install_github("BobyIwan/saeHB.Spatial.Beta")
```

Or, to include the vignette:

``` r
devtools::install_github("BobyIwan/saeHB.Spatial.Beta", build_vignettes = TRUE)
```

## Example

This is a basic example of using the `betadeff_sar()` function to make an estimate based on synthetic data in this package:

```{r example, results='hide', message=FALSE, warning=FALSE}
library(saeHB.Spatial.Beta)

# Load dataset and proximity matrix
data(databeta)
data(weight_mat)

# Fitting the Spatial SAR model
model_sar_deff <- betadeff_sar(
  formula = y ~ x1 + x2,
  deff = "deff",
  n_i = "n_i",
  proxmat = weight_mat,
  data = databeta
)
```

Extract the mean estimation for the areas:

```{r}
head(model_sar_deff$est)
```

Extract the estimated model coefficients:

```{r}
model_sar_deff$coefficient
```

Extract the random effect for the areas:

```{r}
model_sar_deff$randeff
```

Extract the random effect variance for the areas:

```{r}
model_sar_deff$refvar
```

## References
* Rao, J. N. K., & Molina, I. (2015). Small Area Estimation (2nd Edition). New Jersey: John Wiley and Sons, Inc. <doi:10.1002/9781118735855>.
* Kubacki, J., & Jedrzejczak, A. (2016). Small Area Estimation of Income Under Spatial SAR Model. Statistics in Transition New Series, Vol. 17, No. 3, pp. 365--390. <doi:10.59170/stattrans-2016-022>.
* Leroux, B. G., Lei, X., & Breslow, N. (2000). Estimation of Disease Rates in Small Areas: A New Mixed Model for Spatial Dependence. In M. E. Halloran & D. Berry (Eds.), Statistical Models in Epidemiology, the Environment, and Clinical Trials (Vol. 116, pp. 179--191). New York: Springer. <doi:10.1007/978-1-4612-1284-3_4>.
* Chung, H. C., & Datta, G. S. (2020). Bayesian Hierarchical Spatial Models for Small Area Estimation. Research Report Series. Washington, D.C.: U.S. Census Bureau.

