---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# saeHB.Spatial.Beta

The `saeHB.Spatial.Beta` package provides several functions to estimate small area proportions using the Hierarchical Bayesian (HB) method. Model-based estimators are designed for variables of interest that follow a Beta distribution (proportions bounded between 0 and 1). The package supports both non-spatial and spatial models based on Simultaneous Autoregressive (SAR) and Leroux Conditional Autoregressive (CAR) structures for area-level random effects, with optional survey design effect (DEFF) adjustments for sampling variances. In addition, it provides utility functions for constructing spatial weights matrices and performing spatial autocorrelation diagnostics. The `runjags` package is used to obtain posterior estimates via Markov Chain Monte Carlo (MCMC) with parallel computing capabilities. 

## Author
Boby Iwan, Cucu Sumarni

## Maintainer
Boby Iwan <bobyiwanboby2122@gmail.com>

## Functions
* `betadeff_sar()` Estimates small area proportions using a Hierarchical Bayes (HB) method under a Spatial SAR Model with a Beta distribution, incorporating survey design effect (DEFF) adjustments.
* `beta_sar()` Estimates small area proportions using a Hierarchical Bayes (HB) method under a Spatial SAR Model with a Beta distribution without DEFF adjustments, by estimating the unknown precision parameter.
* `betadeff_lerouxcar()` Estimates small area proportions using a Hierarchical Bayes (HB) method under a Spatial Leroux CAR Model with a Beta distribution, incorporating survey design effect (DEFF) adjustments.
* `beta_lerouxcar()` Estimates small area proportions using a Hierarchical Bayes (HB) method under a Spatial Leroux CAR Model with a Beta distribution without DEFF adjustments, by estimating the unknown precision parameter.
* `betadeff_nonspatial()` Estimates small area proportions using a Hierarchical Bayes (HB) method under a Non-Spatial Model with a Beta distribution and Independent and Identically Distributed (IID) random effects, incorporating DEFF adjustments.
* `beta_nonspatial()` Estimates small area proportions using a Hierarchical Bayes (HB) method under a Non-Spatial Model with a Beta distribution and IID random effects without DEFF adjustments, by estimating the unknown precision parameter.
* `build_w()` A utility function to construct spatial weights matrices (contiguity, distance, or kernel) required for spatial modeling.
* `moran_test()` A diagnostic function to perform Moran's I test for spatial autocorrelation.


## Installation
**System Requirement:** Since this package relies on `runjags` for parallel MCMC computations, you must first install the [JAGS (Just Another Gibbs Sampler)](https://mcmc-jags.sourceforge.io/) software on your computer before installing this package.

You can install the development version of saeHB.Spatial.Beta from [GitHub](https://github.com/BobyIwan/saeHB.Spatial.Beta) with:

``` r
# install.packages("devtools")
devtools::install_github("BobyIwan/saeHB.Spatial.Beta")
```

Or, to include the vignette, use the following command:

``` r
devtools::install_github("BobyIwan/saeHB.Spatial.Beta", build_vignettes = TRUE)
```

## Example

This is a basic example of using the `betadeff_sar()` function to make an estimate based on synthetic data in this package:

```{r example, results='hide', message=FALSE, warning=FALSE}
library(saeHB.Spatial.Beta)

# Load dataset and proximity matrix
data(databeta)
data(weight_mat)

# Fitting the Spatial SAR model
model_sar_deff <- betadeff_sar(
  formula = y ~ x1 + x2,
  deff = "deff",
  n_i = "n_i",
  proxmat = weight_mat,
  data = databeta
)
```

Extract the mean estimation for the areas:

```{r}
head(model_sar_deff$est)
```

Extract the estimated model coefficients:

```{r}
model_sar_deff$coefficient
```

Extract the random effect for the areas:

```{r}
model_sar_deff$randeff
```

Extract the random effect variance for the areas:

```{r}
model_sar_deff$refvar
```

## References
* Rao, J. N. K., & Molina, I. (2015). *Small Area Estimation* (2nd ed.). New Jersey: John Wiley & Sons, Inc. <doi:10.1002/9781118735855>.
* Liu, B., Lahiri, P., & Kalton, G. (2014). Hierarchical Bayes Modeling of Survey-Weighted Small Area Proportions. *Statistics Canada*. <https://www150.statcan.gc.ca/n1/pub/12-001-x/2014001/article/14030-eng.pdf>.
* Kubacki, J., & Jedrzejczak, A. (2016). Small Area Estimation of Income Under Spatial SAR Model. *Statistics in Transition New Series*, *17*(3), 365--390. <doi:10.59170/stattrans-2016-022>.
* Leroux, B. G., Lei, X., & Breslow, N. (2000). Estimation of Disease Rates in Small Areas: A New Mixed Model for Spatial Dependence. In M. E. Halloran & D. Berry (Eds.), *Statistical Models in Epidemiology, the Environment, and Clinical Trials* (Vol. 116, pp. 179--191). New York: Springer. <doi:10.1007/978-1-4612-1284-3_4>.
* Chung, H. C., & Datta, G. S. (2020). *Bayesian Hierarchical Spatial Models for Small Area Estimation* (Research Report Series). Washington, D.C.: U.S. Census Bureau. <https://www.census.gov/content/dam/Census/library/working-papers/2020/adrm/RRS2020-07.pdf>.
* Figueroa-Zúñiga, J. I., Arellano-Valle, R. B., & Ferrari, S. L. P. (2013). Mixed beta regression: A Bayesian perspective. *Computational Statistics & Data Analysis*, 61, 137-147. <doi:10.1016/j.csda.2012.12.002>.
* Denwood, M. J. (2016). runjags: An R Package Providing Interface Utilities, Model Templates, Parallel Computing Methods and Additional Distributions for MCMC Models in JAGS. *Journal of Statistical Software*, 71(9), 1–25. <doi:10.18637/jss.v071.i09>.
* Anselin, L. (1988). *Spatial Econometrics: Methods and Models*. Dordrecht: Springer Netherlands. <doi:10.1007/978-94-015-7799-1>.
* Anselin, L., & Morrison, S. (2019). *Spatial Weights as Distance Functions*. <https://spatialanalysis.github.io/lab_tutorials/Spatial_Weights_as_Distance_Functions.html>.
