---
title: "Getting Started with saeHB.Spatial.Beta"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with saeHB.Spatial.Beta}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Introduction

The `saeHB.Spatial.Beta` package provides several functions to estimate small area proportions using Hierarchical Bayesian (HB) methods under spatial and non-spatial models. This package is specifically designed to accommodate survey design effects (DEFF) and handle non-sampled areas.

In this vignette, we will demonstrate a complete analytical workflow:

1. Fitting a Non-Spatial Beta Model.
2. Checking spatial autocorrelation on the random effects using two approaches of Moran's I (Analytical and Monte Carlo).
3. Fitting a Spatial SAR Beta Model.
4. Comparing the precision using Relative Standard Error (RSE) or Coefficient of Variation (CV).

## Step 1: Preparation and Data Loading

First, load the package along with the provided synthetic dataset (`dataBeta`) and proximity matrix (`weight_mat`). We will also load `ggplot2` for visualization.

```{r setup}
library(saeHB.Spatial.Beta)
library(ggplot2)

# Load data and weight matrix
data("dataBeta")
data("weight_mat")

# Calculate Variance of Direct Estimator for proportion data considering DEFF
# var(y) = [y * (1 - y) / n_i] * deff
var_direct <- (dataBeta$y * (1 - dataBeta$y) / dataBeta$n_i) * dataBeta$deff

# Calculate Relative Standard Error (RSE) or CV of Direct Estimation
dataBeta$rse_direct <- (sqrt(var_direct) / dataBeta$y) * 100
```

## Step 2: Fitting the Non-Spatial Model

We begin by fitting a baseline non-spatial model that accommodates the survey design effect. Here, we use the default Markov Chain Monte Carlo (MCMC) iterations.

```{r ns-model, results='hide'}
mod_ns_deff <- betaDeffNonSpatial(
  formula = y ~ x1 + x2,
  DEFF = "deff",
  n_i = "n_i",
  data = dataBeta
)
```

Extract the RSE (or CV) for the non-spatial estimates using the exact column names (Estimate and Est.Error):

```{r ns-rse}
rse_ns_deff <- (mod_ns_deff$Est$Est.Error / mod_ns_deff$Est$Estimate) * 100
```

## Step 3: Spatial Autocorrelation Diagnostic (Moran's I)

To determine if a spatial model is necessary, we evaluate the spatial autocorrelation of the random effects ($v$) from the non-spatial model. For robust validation, we use both the Analytical approach and the Monte Carlo permutation approach.

```{r moran-test}
# Convert the proximity matrix to a listw object
W_listw <- spdep::mat2listw(weight_mat, style = "W")

# Extract the mean of the random effects (v) using the exact column name 'Estimate'
v_ns_deff <- as.numeric(mod_ns_deff$randeff$Estimate)

# 1. Analytical Approach
moran_analytical <- spatial_moran(x = v_ns_deff, listw = W_listw)
print(moran_analytical)

# 2. Monte Carlo Permutation Approach
set.seed(123) # For reproducibility
moran_mc <- spdep::moran.mc(x = v_ns_deff, listw = W_listw, nsim = 999)
print(moran_mc)
```

A significant p-value in both tests confirms that the non-spatial model left unexplained spatial structure in the random effects, heavily justifying the use of a spatial model.

## Step 4: Fitting the Spatial SAR Model

Proceeding with the Spatial Simultaneous Autoregressive (SAR) model to capture the spatial dependency:

```{r sar-model, results='hide'}
mod_sar_deff <- betaDeffSAR(
  formula = y ~ x1 + x2,
  DEFF = "deff",
  n_i = "n_i",
  proxmat = weight_mat,
  data = dataBeta
)
```

Extract the RSE (or CV) for the SAR estimates using the exact column names:

```{r sar-rse}
rse_sar_deff <- (mod_sar_deff$Est$Est.Error / mod_sar_deff$Est$Estimate) * 100
```

## Step 5: Evaluation and Comparison

We compare the performance of the Direct Estimator, the Non-Spatial Model, and the Spatial SAR Model. A lower RSE (CV) indicates a more reliable and precise estimate.

```{r comparison, fig.width=7, fig.height=5}
# Combine RSEs into a single data frame for plotting
df_rse <- data.frame(
  Area = seq_along(dataBeta$y),
  Direct = dataBeta$rse_direct,
  Non_Spatial = rse_ns_deff,
  Spatial_SAR = rse_sar_deff
)

# Order by Direct RSE for better visualization
df_rse <- df_rse[order(df_rse$Direct), ]
df_rse$Area_Index <- seq_len(nrow(df_rse))

# Plotting the RSE (CV) Comparison
ggplot(df_rse, aes(x = Area_Index)) +
  geom_line(aes(y = Direct, color = "Direct Estimation"), linewidth = 1, alpha = 0.6) +
  geom_line(aes(y = Non_Spatial, color = "HB Beta Deff Non-Spatial"), linewidth = 1, alpha = 0.8) +
  geom_line(aes(y = Spatial_SAR, color = "HB Beta Deff Spatial SAR"), linewidth = 1.2) +
  scale_color_manual(
    name = "Estimator",
    values = c("Direct Estimation" = "#E69F00", 
               "HB Beta Deff Non-Spatial" = "#56B4E9", 
               "HB Beta Deff Spatial SAR" = "#009E73")
  ) +
  labs(
    title = "Comparison of Relative Standard Error (RSE / CV)",
    subtitle = "Lower RSE indicates higher precision",
    x = "Area (Ordered by Direct RSE)",
    y = "RSE (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

The plot above clearly illustrates how model-based estimators, particularly the spatial ones, shrink the variance and improve the overall precision compared to the direct survey estimates.
